{\fontfamily{ptm}\selectfont A Proximal Point framework of Nesterov type accelerated gradient}

### 2024-08-06 Notebook LaTEX/Momentum Accelerations/main.tex
>w<, a lot of new updates

\title{{\fontfamily{ptm}\selectfont Nesterov Type Momentum Methods}}

@@

            \begin{definition}[Proximal Gradient Tri-Points Form E]
            \label{def:ag_tri_pt_form_E}
                Form E is a special case of 
                \hyperref[def:ag_prox_grad_tri_pt]{definition \ref*{def:ag_prox_grad_tri_pt}}
                with the condition that $\tilde \eta_t = \eta_t$ for all $t \in \mathbb N$. 
                As a consequence, the algorithm has the following updates for iterates $(y_t, x_{t + 1}, z_{t + 1})$: 
                \begin{align*}
                    y_t &= (1 + L\eta_t)^{-1}(x_t + L\eta_t z_t)
                    \\
                    x_{t + 1} &= x_t - \eta_t \mathcal G_L(y_t)
                    \\
                    z_{t + 1} &= y_t - L^{-1}\mathcal G_L(y_t).
                \end{align*}
            \end{definition}

            \begin{definition}[Similar Triangle Form I]
            \label{def:ag_form_similar_tria_I}
                It updates iterates $(y_t, x_{t + 1}, z_{t + 1})$ using 
                \begin{align*}
                    z_{t + 1} &= y_t - L^{-1} \mathcal G_L(y_t) 
                    \\
                    x_{t + 1} &= z_{t + 1} + L\eta_t (z_{t + 1} - z_t)
                    \\
                    y_{t + 1} &= 
                    (1 + L\eta_{t + 1})^{-1}
                    (x_{t + 1} + L\eta_{t + 1}z_{t + 1}). 
                \end{align*}
            \end{definition}
            \begin{observation}
                There are a lot of observations about this form to unpack. 
                Let's list the following observations
                \begin{enumerate}
                    \item Observe that $y_t = (1 + L\eta_t)^{-1}(x_t + L\eta_t z_t)$ is a convex combination between $x_t, z_t$. Hence $z_t, y_t, x_t$ are three collinear points. Additionally, we have $y_t - x_t = L\eta_t (z_t - y_t)$, hence $\Vert y_t - x_t\Vert/\Vert z_t - y_t\Vert = L\eta_t$. 
                    \item By the algorithm, we have $z_{t + 1} - y_t = - L^{-1} \mathcal G_L(y_t)$, and $x_{t + 1} - x_t = - \tilde \eta_{t + 1} \mathcal G_L(y_t)$, hence vector $z_{t + 1} - y_t$ parallels to $x_{t + 1} - x_t$. 
                    \item Finally, with update $x_{t +1} - z_{t + 1} = L\eta_t (z_{t +1} - z_t)$, we have three colinear points $(z_t, z_{t + 1}, x_t)$ with $\Vert z_{t + 1} - z_t\Vert/\Vert z_{t +1 } - z_t\Vert = L\eta_t$. 
                \end{enumerate}
                From the above results, we can conclude that triangle $(y_t, z_t, z_{t + 1})$ is similar to $(x_t, z_t, x_{t + 1})$ because they share colinear points $z_t, y_t, x_t$ and $z_t, z_{t + 1}, x_{t + 1}$, and their sides $z_{t + 1} - y_t$ parallels to $x_{t + 1} - x_t$. 
            \end{observation}
            \begin{remark}
                For more information about the specific choice of $\eta_t$ allows for an optimal convergence rates, see 
                \hyperref[sec:scenario_ii]{section \ref*{sec:scenario_ii}} for more information. 
            \end{remark}

@@

    \subsection{Generic Lyapunov analysis for Accelerated gradient via PPM}
    \label{sec:generic_ag_ppm_lyapunov_analysis}
        The first theorem states the generic convergence results when $f$ is differentiable with $L$-Lipschitz gradient, using \hyperref[def:agg_ppm]{definition \ref*{def:agg_ppm}}.

@@

            we derive the non-smooth analogous case of the Lyapunov upper bound: 

@@

        \begin{theorem}[Generic AG Convergence]
        \label{thm:generic_ag_convergence}
            If there is a choice of $\eta_i, \tilde \eta_i$ in \hyperref[def:agg_ppm]{definition \ref*{def:agg_ppm}} such that we have 
            \begin{align*}
                \left(
                    \sum_{i = 1}^{t} \tilde \eta_i
                \right)
                \Upsilon_{2, t + 1}^{\text{AG}} + 
                \Upsilon_{1, t + 1}^{\text{AG}} \le 0, 
            \end{align*}
            then with 
            \begin{align*}
                \Phi_t &= \left(
                    \sum_{i = 1}^{t} \tilde\eta_{i}
                \right) (f(z_t) - f(x_*)) + \frac{1}{2}\Vert x_t - x_*\Vert^2 \quad \forall t \in \NN
                \\
                \Phi_0 &= \frac{1}{2}\Vert x_t - x_*\Vert^2, 
            \end{align*}
            we have $\Phi_{t + 1} - \Phi_t \le 0$, which allows for a convergence rate of $\mathcal O \left(\sum_{i = 1}^{T} \eta_i^{-1}\right)$ for $f(z_T) - f(x_*)$. 
        \end{theorem}
        \begin{proof}
            By definition we have
            {\footnotesize
            \begin{align*}
                \Phi_{t + 1} - \Phi_t 
                &= 
                \left(
                    \sum_{i = 1}^{t+1} \tilde\eta_{i}
                \right) (f(z_{t + 1}) - f(x_*)) 
                - 
                \left(
                    \sum_{i = 1}^{t} \tilde\eta_{i}
                \right) (f(z_{t}) - f(x_*)) 
                + \frac{1}{2}\Vert x_t - x_*\Vert^2
                - \frac{1}{2}\Vert x_{t + 1} - x_*\Vert^2
                \\
                &= 
                \tilde \eta_{t + 1} (f(z_{t + 1}) - f(z_*))
                +
                \left(
                    \sum_{i = 1}^{t} \tilde \eta_i
                \right)(f(z_{t + 1}) - f(z_t))
                + \frac{1}{2}\Vert x_t - x_*\Vert^2
                - \frac{1}{2}\Vert x_{t + 1} - x_*\Vert^2
                \\
                &= \left(
                    \sum_{i = 1}^{t} \tilde \eta_i
                \right)\Upsilon_{2, t + 1}^{\text{AG}} + \Upsilon_{1, t + 1}^{\text{AG}} \le 0. 
            \end{align*}
            }
            The derivation for the convergence raet is direct and similar to 
            \hyperref[thm:ppm_convergence_rate]{theorem \ref*{thm:ppm_convergence_rate}} 
            Therefore, if we can identify parameter for the generic algorithm that asserts the condition above, then we have convergence for the algorithm. 
        \end{proof}

>>>>

\title{{\fontfamily{ptm}\selectfont A Proximal Point framework of Nesterov type accelerated gradient}}

@@

            \begin{definition}[Proximal Gradient variant of Ryu 12.1]
            \label{def:ag_tri_pt_form_E}
                The method appeared in Ryu's text is a special case of 
                \hyperref[def:ag_prox_grad_tri_pt]{definition \ref*{def:ag_prox_grad_tri_pt}}
                with the condition that $\tilde \eta_t = \eta_t$ for all $t \in \mathbb N$ and $g \equiv 0$ for the non-smooth part of the objective. 
                As a consequence, the algorithm has the following updates for iterates $(y_t, x_{t + 1}, z_{t + 1})$: 
                \begin{align*}
                    y_t &= (1 + L\eta_t)^{-1}(x_t + L\eta_t z_t)
                    \\
                    x_{t + 1} &= x_t - \eta_t \mathcal G_L(y_t)
                    \\
                    z_{t + 1} &= y_t - L^{-1}\mathcal G_L(y_t).
                \end{align*}
            \end{definition}

            \begin{definition}[Similar Triangle Form I]
            \label{def:ag_form_similar_tria_I}
                The similar triangle form is equivalent to the classic nesterov accelerated gradient proposed back in 1983. 
                It updates iterates $(y_t, x_{t + 1}, z_{t + 1})$ using 
                \begin{align*}
                    z_{t + 1} &= y_t - L^{-1} \mathcal G_L(y_t) 
                    \\
                    x_{t + 1} &= z_{t + 1} + L\eta_t (z_{t + 1} - z_t)
                    \\
                    y_{t + 1} &= 
                    (1 + L\eta_{t + 1})^{-1}
                    (x_{t + 1} + L\eta_{t + 1}z_{t + 1}). 
                \end{align*}
            \end{definition}
            \begin{observation}
                There are a lot of observations about this form to unpack. 
                Let's list the following observations
                \begin{enumerate}
                    \item Observe that $y_t = (1 + L\eta_t)^{-1}(x_t + L\eta_t z_t)$ is a convex combination between $x_t, z_t$. Hence $z_t, y_t, x_t$ are three collinear points. Additionally, we have $y_t - x_t = L\eta_t (z_t - y_t)$, hence $\Vert y_t - x_t\Vert/\Vert z_t - y_t\Vert = L\eta_t$. 
                    \item By the algorithm, we have $z_{t + 1} - y_t = - L^{-1} \mathcal G_L(y_t)$, and $x_{t + 1} - x_t = - \tilde \eta_{t + 1} \mathcal G_L(y_t)$, hence vector $z_{t + 1} - y_t$ parallels to $x_{t + 1} - x_t$. 
                    \item Finally, with update $x_{t +1} - z_{t + 1} = L\eta_t (z_{t +1} - z_t)$, we have three colinear points $(z_t, z_{t + 1}, x_t)$ with $\Vert z_{t + 1} - z_t\Vert/\Vert z_{t +1 } - z_t\Vert = L\eta_t$. 
                \end{enumerate}
                From the above results, we can conclude that triangle $(y_t, z_t, z_{t + 1})$ is similar to $(x_t, z_t, x_{t + 1})$ because they share colinear points $z_t, y_t, x_t$ and $z_t, z_{t + 1}, x_{t + 1}$, and their sides $z_{t + 1} - y_t$ parallels to $x_{t + 1} - x_t$. 
            \end{observation}
            \begin{remark}
                The above algorithm is a special case of 
                \hyperref[def:ag_prox_grad_tri_pt]{definition \ref*{def:ag_prox_grad_tri_pt}}, 
                see 
                \hyperref[sec:scenario_ii]{section \ref*{sec:scenario_ii}} for more information. 
            \end{remark}

@@

    \subsection{Generic Lyapunov analysis for Accelerated gradient via PPM}
    \label{sec:generic_ag_ppm_lyapunov_analysis}
        The lemma theorem states targets
        \hyperref[prop:derive_ag_prox_grad_tript]
        {definition \ref*{prop:derive_ag_prox_grad_tript}}.

@@

            we derive the non-smooth analogous case of the Lyapunov upper bound for all $t \in \mathbb N \cup \{0\}$: 

@@

        \begin{theorem}[Generic AG Convergence]
        \label{thm:generic_ag_convergence}
            If there is a choice of $\eta_i, \tilde \eta_i$ in \hyperref[def:agg_ppm]{definition \ref*{def:agg_ppm}} such that we have 
            \begin{align*}
                \left(
                    \sum_{i = 1}^{t} \tilde \eta_i
                \right)
                \Upsilon_{2, t + 1}^{\text{AG}} + 
                \Upsilon_{1, t + 1}^{\text{AG}} 
                &\le \delta_{t + 1} \quad 
                \forall t \in \mathbb N, 
                \\
                \Upsilon_{1, 1}^{\text{AG}} &\le \delta_1. 
            \end{align*}
            then with 
            \begin{align*}
                S_t &:= \sum_{i = 1}^{t} \delta_i
                \\
                \sigma_t &:= \sum_{i = 1}^{t}\tilde \eta_i
                \\
                \Phi_t &= \left(
                    \sum_{i = 1}^{t} \tilde\eta_{i}
                \right) (f(z_t) - f(x_*)) + \frac{1}{2}\Vert x_t - x_*\Vert^2 \quad \forall t \in \NN
                \\
                \Phi_0 &= \frac{1}{2}\Vert x_0 - x_*\Vert^2, 
            \end{align*}
            we have $\Phi_{t + 1} - \Phi_t \le \delta_{t + 1}$, which allows for a convergence rate of $\mathcal O \left(\sigma_T^{-1}S_T\right)$ of $f(z_T) - f(x_*)$ when $S_T \neq 0$ and a rate of $\mathcal O(\sigma_T^{-1})$ when $S_T = 0$. 
        \end{theorem}
        \begin{proof}
            By definition we have
            {\footnotesize
            \begin{align*}
                \Phi_{t + 1} - \Phi_t 
                &= 
                \left(
                    \sum_{i = 1}^{t+1} \tilde\eta_{i}
                \right) (f(z_{t + 1}) - f(x_*)) 
                - 
                \left(
                    \sum_{i = 1}^{t} \tilde\eta_{i}
                \right) (f(z_{t}) - f(x_*)) 
                + \frac{1}{2}\Vert x_t - x_*\Vert^2
                - \frac{1}{2}\Vert x_{t + 1} - x_*\Vert^2
                \\
                &= 
                \tilde \eta_{t + 1} (f(z_{t + 1}) - f(z_*))
                +
                \left(
                    \sum_{i = 1}^{t} \tilde \eta_i
                \right)(f(z_{t + 1}) - f(z_t))
                + \frac{1}{2}\Vert x_t - x_*\Vert^2
                - \frac{1}{2}\Vert x_{t + 1} - x_*\Vert^2
                \\
                &= \left(
                    \sum_{i = 1}^{t} \tilde \eta_i
                \right)\Upsilon_{2, t + 1}^{\text{AG}} + \Upsilon_{1, t + 1}^{\text{AG}} \le \delta_{t + 1}. 
            \end{align*}
            }
            Telescoping we have 
            \begin{align*}
                \Phi_{t + 1} - \Phi_t &\le \delta_{t + 1}
                \\
                \sum_{i = 0}^{T - 1}\Phi_{i + 1} - \Phi_i &\le \sum_{i = 0}^{T - 1}\delta_{i + 1}
                \\
                \Phi_T - \Phi_0 &\le 
                \sum_{i = 1}^{T}\delta_i = S_{T}. 
            \end{align*}
            So then $\Phi_T - \Phi_0$ yields: 
            $$
            \begin{aligned}
                \sigma_T (f(z_t) - f(x_*)) 
                + \frac{1}{2}\Vert x_t - x_*\Vert^2 
                - \frac{1}{2}\Vert x_0 - x_*\Vert^2 
                &\le S_{T}
                \\
                \implies 
                \sigma_T(f(z_t) - f(x_*))
                &\le 
                S_T + \frac{1}{2}\Vert x_0 - x_*\Vert^2
                \\
                f(z_t) - f(x_*) &\le 
                \sigma_T^{-1}\left(
                    S_{T} + \frac{1}{2}\Vert x_0 - x_*\Vert^2
                \right),
            \end{aligned}
            $$
            which yields a convergence rate $\mathcal O(\sigma_T^{-1}S_{T})$. 
            When $S_T = 0$, the convergence rate is $O(\sigma_T^{-1})$ instead. 
        \end{proof}

        \begin{theorem}[Constraints on PPM stepsizes]
            Algorithm proposed back in 
            \hyperref[def:ag_prox_grad_tri_pt]{definition \ref*{def:ag_prox_grad_tri_pt}}
            converges if the stepsize parameter. 
        \end{theorem}
